### 2025-01-23 paper/related.tex
bibliography sorted and everything adapted to ECML

@@

generate counterfactual images for MNIST and ImageNet through
independent mechanisms (IM): each IM learns class-conditional input
distributions over a specific lower-dimensional, semantically meaningful
factor, such as \emph{texture}, \emph{shape} and \emph{background}. The
demonstrate that using these generated counterfactuals during classifier
training improves model robustness. Similarly, argue that
counterfactuals represent potentially useful training data in machine
learning, especially in supervised settings where inputs may be
reasonably mapped to multiple outputs. They, too, demonstrate the
augmenting the training data of image classifiers can improve
generalization.

propose an approach using counterfactuals in training that does not rely
on data augmentation: they argue that counterfactual pairs typically
already exist in training datasets. Specifically, their approach relies
on, firstly, identifying similar input samples with different
annotations and, secondly, ensuring that the gradient of the classifier
aligns with the vector between pairs of counterfactual inputs using the
cosine distance as a loss function (referred to as \emph{gradient
supervision})

@@

propose two definitional differences between Adversarial Examples (AE)
and Counterfactual Explanations (CE): firstly, and more importantly
according to the authors, the term AE implies missclassification, which
is not the case for CE;

@@

secondly, they argue that closeness plays a more critical role in the
context of CE but confess that even counterfactuals that are not close
might be relevant explanations. show that CE and AE are equivalent under
certain conditions and derive upper bounds on the distances between
them.

@@

are the first to propose end-to-end training pipeline that includes
counterfactual explanations as part of the training prodeduce. In
particular, they propose a specific network architecture that includes a
predictor and CE generator network,

@@

propose a way to train models that are guaranteed to provide recourse
for individuals with high probability. The approach builds on
adversarial training,

@@

introduce Counterfactual Adversarial Training (CAT) with intention of
improving generalization and robustness of language models.
Specifically, they propose to proceed as follows: firstly, identify
training samples that are subject to high predictive uncertainty
(entropy); secondly, generate counterfactual explanations for those
samples; and, finally, finetune the model on the augmented dataset that
includes the generated counterfactuals.

>>>>

\cite{wachter2017counterfactual, joshi2019realistic, altmeyer2024faithful}

@@

\cite{sauer2021counterfactual} generate counterfactual images for MNIST
and ImageNet through independent mechanisms (IM): each IM learns
class-conditional input distributions over a specific lower-dimensional,
semantically meaningful factor, such as \emph{texture}, \emph{shape} and
\emph{background}. The demonstrate that using these generated
counterfactuals during classifier training improves model robustness.
Similarly, \cite{abbasnejad2020counterfactual} argue that
counterfactuals represent potentially useful training data in machine
learning, especially in supervised settings where inputs may be
reasonably mapped to multiple outputs. They, too, demonstrate the
augmenting the training data of image classifiers can improve
generalization.

\cite{teney2020learning} propose an approach using counterfactuals in
training that does not rely on data augmentation: they argue that
counterfactual pairs typically already exist in training datasets.
Specifically, their approach relies on, firstly, identifying similar
input samples with different annotations and, secondly, ensuring that
the gradient of the classifier aligns with the vector between pairs of
counterfactual inputs using the cosine distance as a loss function
(referred to as \emph{gradient supervision})

@@

\cite{freiesleben2022intriguing} propose two definitional differences
between Adversarial Examples (AE) and Counterfactual Explanations (CE):
firstly, and more importantly according to the authors, the term AE
implies missclassification, which is not the case for CE;

@@

secondly, they argue that closeness plays a more critical role in the
context of CE but confess that even counterfactuals that are not close
might be relevant explanations. \cite{pawelczyk2022exploring} show that
CE and AE are equivalent under certain conditions and derive upper
bounds on the distances between them.

@@

\cite{guo2023counternet} are the first to propose end-to-end training
pipeline that includes counterfactual explanations as part of the
training prodeduce. In particular, they propose a specific network
architecture that includes a predictor and CE generator network,

@@

\cite{ross2021learning} propose a way to train models that are
guaranteed to provide recourse for individuals with high probability.
The approach builds on adversarial training,

@@

\cite{luu2023counterfactual} introduce Counterfactual Adversarial
Training (CAT) with intention of improving generalization and robustness
of language models. Specifically, they propose to proceed as follows:
firstly, identify training samples that are subject to high predictive
uncertainty (entropy); secondly, generate counterfactual explanations
for those samples; and, finally, finetune the model on the augmented
dataset that includes the generated counterfactuals.
